\documentclass[11pt]{article}
\bibliographystyle{siam}

\title{The title of your project proposal}
\author{
  Daks, Alon\\
  \texttt{alondaks}
  \and
  Yu, Lisa Ann\\
  \texttt{lisaannyu}
  \and
  Luo, Ying\\
  \texttt{yingtluo}
  \and
  Chang, Jordeen\\
  \texttt{Jodreen}
}

\begin{document}
\maketitle

Identify a published fMRI paper and the accompanying data
\cite{lindquist2008statistical}.  You should explain the basic idea of the
paper in a paragraph.  You should also perform basic sanity check on the data
(e.g., can you downloaded, can you load the files, confirm that you have the
correct number of subjects).

Hanke, Michael, et al. "A high-resolution 7-Tesla fMRI dataset from complex 
natural stimulation with an audio movie." Scientific data 1 (2014).

Most fMRI studies use highly simplified stimulations that are vastly dissimilar
from what people experience in everyday life.  Hanke et al. sought to create a
dataset of naturally occurring brain states by exposing participants to a 
more complex stimulus, the audio description of Forrest Gump.  This particular
audio description allows for the study of auditory attention and cognition, 
language and music perception, and retrieval of explicit memory without the
effect of visual imagery.  In addition, this dataset uses inter-individual 
synchronicity to study responses to complex processing.  This paper explains
the preparation of the dataset, such as collecting physiological information
then accounting for it in the voxels, and an initial analysis.  Hanke et al.
used representational similarity analysis to identify similar patterns
across brains.  They created a representationaly consistency map using pairwise
correlations.  After thresholding, the map revealed two clusters in the PT,
which is known to be used for auditory cognition.

The main purpose of this experiment was to study properties of brain response
patterns that are supposedly common when people are exposed to audio and movie
simulation. We intend to replicate their experiment using the data they gathered
from the 20 participants. For example, a BOLD time-series similarity measure
(e.g. correlation) is often used to quantify similarities in responses among
individuals. Hank et al. recognized that this was a common approach, but they
went beyond that and also implemented representational similarity analysis (RSA).
To do so, we will create dissimilarity matrices for 18 individuals using the same
searchlight mapping approach that they used (Subjects 4 and 10 were not included
due to missing data). Doing so will capture 2nd-order isomorphisms in response
patterns. Lastly, to access statistical significance, we will transform the
representational consistency map into percent rank with respect to the total
distribution of the DSM correlations. We'll calculate the mean correlation
coefficient and compare our value to theirs.



\bibliography{project}

\end{document}
